{Simulating the Brownian Dynamics of a Two-dimensional Model for the Dynein Motor Protein}\vspace{1.5em}\\ \includegraphics[width=0.5\columnwidth]{OSU_logo.png}\vspace{0.5em}\\ \large{Oregon State University}

### 2019-05-07 papers/waczak-thesis/thesis/chapters/methods.tex

\section{Testing the model} 
\subsection{Sanity Checks}
This sub section will be for the various methods we used to ensure our model is behaving reasonably 
\subsubsection{Verifying size of time step} 
Here I will talk about how we found autocorrelations functions for each domain energy at various values of dt in order to verify our time step is as large as can be
\subsubsection{Energy conservation and the Equipartion theorem} 
talk about code to check that we satisfy energy 
\subsection{Fitting to other studies}

>>>>

\section{The Code}
The simulation code was written in C++ in order to take advantage of the language's speed and object-oriented programming features. The equations of motion for our model in each state form a linear system of coupled first order differential equations for the domain velocities which we solved directly using Mathematica. We employed git version control to log changes to the code base-- our digital lab notebook. Data sets were stored locally and data used for our analysis was added to an online repository. An up-to-date version can be found at \url{https://github.com/droundy/dynein_walk}.\\

Armed with equations for each domain velocity, we can now describe how the simulation evolves our model through time, that is, how we integrate for the time evolution of each state \textit{and} how we negotiate binding and unbinding events.

	\section{Time evolution: Euler's method}
	There are countless numerical methods for integrating differential equations. Popular methods for molecular dynamics simulations typically include Verlet integration or $n^{th}$ order Runge-Kutta. The first is valuable as it preserves time-reversal symmetry and the second integrates to a high level of precision. These properties are nice, but the randomness of Brownian motion means that there is little sense in running a simulation backwards and the degree of extra precision afforded is meaningless. We therefore employ perhaps the simplest integration scheme: Euler's method. \\

	Euler's method assumes that a function $f$ can be treated as locally linear so that the tangent line is a good approximation for the value of the function. Therefore, if we know the value of $f$ at some time $t$, then the value of the function at a nearby time $t+\Delta t$ can be approximated by 
	\begin{equation}
		f(t+\Delta t) = f(t)+f'(t)\Delta t
	\end{equation}
	where $\Delta t$ is understood to be small. We apply this scheme to the system of equations given by equation (3.15) to adjust the positions of each domain and \textit{move} the protein.

	\subsection{The Random Force}
	As noted in section 3.1, dynein experiences a frequent, random interaction force due to collisions between each domain and the surrounding water molecules. To simulate this extra force, we sample from a zero centered Gaussian distribution for each domain during each time step.

	\subsection{Putting it all together}
	The last piece of the simulation is deciding when dynein has binds to or unbinds from the microtubule. This is accomplished via turning the binding (eq 3.26) and unbinding rates (eq 3.27) into probabilities, e.g. 
	\begin{align}
		P_{\text{bind}} &= \rho_b\; \Delta t \\ 
		P_{\text{unbind}} &= \rho_{ub}\;\Delta t
	\end{align}

	During each time step, we first update the velocity of each domain and then use equation (4.1) to evolve their positions. Next, we calculate the binding/unbinding probabilities for equations (4.2) and (4.3). Finally, we \textit{roll a die} to select a random number in the interval $[0,1]$. If this number is below the binding/unbinding probability, we record the step and then change dynein's state. If it is above, we repeat the process and time evolve the protein again. \\

	This process is outlined for the one bound state in the following flowchart.

	\begin{figure}[!hbt]
		\centering
		\includegraphics[width=0.39\columnwidth]{One_bound_sim_flowchart.pdf}
		\caption[One bound simulation flow chart]{\textbf{One bound simulation flow chart} showing how the simulation is performed for the one bound state. Note that we do not allow dynein to pass through the microtubule as this is a 2-dimensional simulation.} 
	\end{figure}

	 \newpage

\section{Determining constants of the model}
In order to determine the stalk and tail lengths $\ell_S, \ell_T$, domain radii, and the set of equilibrium angles $\{\varphi_{i,\text{eq}}\}$, we examined microscopy images of dynein in various configurations. The following figure shows a cartoon of the model superimposed on images of real dynein in what we identify as the both bound and one bound states. 
\begin{figure}[hbt!]
	\centering
	\includegraphics[width=0.3\columnwidth]{burgess-model-figure.pdf}
	\includegraphics[width=0.5\columnwidth]{grotjahn-model-figure.pdf}%
	\caption[Model parameters determined from experiment ]{\textbf{Model parameters determined from experiment} Left: the both bound model is superimposed over dynein micrograph. The arrow indicates 15nm. Image taken from \cite{burgess2003dynein}.  Right: cryo-electron tomograph of dynein with the one-bound model overlaid. Image taken from \cite{grotjahn}.} 
	\label{fig:superimpmosed}
\end{figure}
Figure \ref{fig:superimpmosed} shows example images of micrographs used to determine the model constants. In order to account for the fact that we idealize the stalks and tail as thin, massless rods, we slightly increase the size of each domain to account for any extra drag. Interestingly, these images suggest an angle of $0^\circ$ for both states.

\section{Testing the model} 
To test the model we ran multiple simulations varying the spring constants and binding / unbinding rates. Having so many degrees of freedom makes using any kind of minimization process challenging to implement and far too long to simulate. Nevertheless, we can still be clever by using reported velocities (for example \cite{dewitt2012cytoplasmic} reports 124 nm/sec) to estimate the time spent in the one bound and both bound states. With this information we can set the binding and unbinding rates in such a way as to simulate only the both bound state \textit{or} the one bound state. With these simulations we can get good estimates for the rate constants.

With good estimates for these constants, we then refine the values by fitting our stepping statistics to those of experiments such as shown in figures \ref{fig:weihong histogram} and \ref{fig:yildiz_tension}.

### 2019-05-07 papers/waczak-thesis/thesis/chapters/results.tex
big thesis update 1
thesis update 2
thesis update

\todo[inline]{TODO: I plan to add description and section headings over break}
\todo[inline]{TODO: add discussion of powerstroke vs winch models. Discuss simple modifications to simulation parameters to try and fit for this alternate mechanochemical model}

\todo[inline]{TODO: combine 5.1 with 5.2 to explain leading vs lagging}
\todo[inline]{TODO: consider removing leading lagging axis labels in 5.1 if redundant with figure caption} 
\todo[inline]{NOTE: it looks like there may be an issue with the middle most cartoon in 5.3. The tail does not appear to be the correct size}
\todo[inline]{TODO decrease overall width of tail in cartoon dynein. Increase thickness of stalk} 
\todo[inline]{TODO: Check to make sure time axis in final panel of fig 5.4 is actually a log scale}

\section{Stepping}
\begin{figure}
	\centering
	\includegraphics[width=1\columnwidth]{paper_trajectory_plot}
	\caption{\textbf{Trajectory Plot} Vertical and horizontal projections of each of binding domains. Directed motion is clearly seen in the positive slope of the x-projection graph. Dynein's conformation is illustrated for a selection of times.}
	\label{fig:trajectory}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.85\columnwidth]{paper_model_behavior}
	\caption{\textbf{Stepping statistics} trajectories run with different random number seeds all show forward directed motion. Step lengths have a similar distribution to experiment although steps of zero length are not measured experimentally. }
	\label{Fig:statistics}
\end{figure}

\section{One bound and both bound correlations}
\begin{figure}[!hbt]
	\centering
	\includegraphics[width=0.75\columnwidth]{paper_displacement_vs_step_length}
	\caption{\textbf{Two dimensional histogram} comparing the total signed step length to
			the signed initial displacement between the binding domains. A positive value
			indicates that the front foot made the step (leading). A negative value
			indicates the rear foot (lagging) made the step. Simulations show a strong,
			negative correlation. A similar fit from an experimental paper is plotted for comparison.}
	\label{fig:displacement_vs_step_length}
\end{figure}

\begin{figure}[!hbt]
	\centering
	\includegraphics[angle=-90, width=1\columnwidth]{leading_vs_lagging}
	\caption{Diagram illustrating a leading foot step (a) and a lagging foot step (b).}
\end{figure}

\section{Possible Monte Carlo}
\begin{figure}[!hbt]
	\centering
	\includegraphics[width=1\columnwidth]{energy_landscape}
	\caption{\textbf{Energy Landscape} graph showing the total energy of the dynein for a given binding domain displacement $L$. The yellow and red dots correspond to the minimum and maximum energy configurations for which a cartoon is shown on the right.}
	\label{fig:energy_landscape}
\end{figure}

>>>>

\section{Best-fit parameters}
After running multiple simulations as discussed in section 4.4, the current set of best parameters for our model are shown in the table below
\begin{table}[!hbt]
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
		\hline
		$\mathbf{k_b}$&  $\mathbf{k_{ub}}$ &  $\mathbf{c_b}$&  $\mathbf{c_m}$& $\mathbf{c_t}$ & $\mathbf{\ell_s}$ & $\mathbf{\ell_t}$ & $\mathbf{r_t}$ & $\mathbf{r_m}$ &$\mathbf{r_b}$ & $\mathbf{dt}$ & $\mathbf{c}$\\  
		$s^{-1}$ & $s^{-1}$ & $k_bT$ & $k_bT$ & $k_bT$& nm & nm & nm & nm & nm & $s$ &  \\
		\hline \hline
		$9.5e9$ & $1e20$ & $0.08$ & $1.05$ & $0.36$ & $20.75$ & $23.0$ & $6.0$ & $5.5$ & $1.75$ & $1e10$ & $0.35$\\ 
		\hline
	\end{tabular}
\caption[Best parameters]{Current set of best fit parameters}
\label{table:params}
\end{table}

With these parameters, the following analysis was performed for multiple simulations with a variety of different random number seeds.

\section{Stepping}

Our model is able to achieve directed motion with clear evidence of the desired Brownian jiggling. All simulations show a net displacement in the positive x direction.
\begin{figure}[!hbt]
	\centering
	\includegraphics[width=1\columnwidth]{paper_trajectory_plot}
	\caption[Trajectory plot]{\textbf{Trajectory Plot} for a small subset of steps. Vertical and horizontal projections for the trajectory of each binding domain are plotted in red and blue. Directed motion is clearly seen in the positive slope of the x-projection graph. Dynein's conformation is illustrated for a selection of times.}
	\label{fig:trajectory}
\end{figure}

As y-projection plot in figure \ref{fig:trajectory} indicates, the model performs a wide diffusive search for the microtubule, assuming a wide range of configurations along the way. The cartoons corresponding to the black triangles in the x and y projection plots show the motor domains travel around the tail linker during the stepping cycle. This behavior is suggestive of the power-stroke model proposed by \cite{burgess2003dynein} and others. It is also worth noting that the y-projection plot suggests an average binding domain height $>0$. This is contrary to the popular depictions of dynein with the binding domains nearly scrubbing the microtubule until binding.\\

Dynein's stepping velocity can be seen in the positive slope of figure \ref{fig:trajectory}. This velocity is directly related to the stepping rates. To increase the model's velocity, simply increase both the binding and unbinding rates. \\

Running a longer simulations at multiple random number seeds allows us to gather more stepping statistics. A sample of these are shown in figure \ref{fig:statistics}.

\begin{figure}
	\centering
	\includegraphics[width=0.85\columnwidth]{paper_model_behavior}
	\caption[Model Stepping Statistics]{\textbf{Stepping statistics} trajectories run with different random number seeds all show forward directed motion. Step lengths have a similar distribution to experiment although steps of zero length are not measured experimentally. }
	\label{fig:statistics}
\end{figure}

The first graph in figure \ref{fig:statistics} shows the x position trajectory plot for the same simulation run with many different random seeds. An experimental trajectory plot is shown in blue with data taken from \cite{dewitt2012cytoplasmic}. This figure illustrates that even for long simulations, the model shows directed motion despite occasionally stepping backwards. Furthermore, we can see that average velocity across each of the model simulations agrees well with the experiment.

The second graph in figure \ref{fig:statistics} shows a signed step length histogram comparing the model with data from \cite{qiu2012dynein}. Here negative values of step length indicate a backwards step. The shape of the histogram matches well with the experiment despite missing the notable dip at zero step length. This is likely due to the fact that a step of zero length (i.e. a stutter step) can not be distinguished experimentally from the typical Brownian motion. \\

Finally, we plot histograms for the time spent in the one bound and both bound states during each simulation. The red line indicates our estimates for these values based off of experimental stepping velocities. It is interesting to note that in both cases, the average time spent in each state is below this estimate. Despite this, we can clearly see that the median one bound time is around $0.0015-0.003$ seconds while the both bound time is between lies somewhere between $0.015-0.03$ seconds-- an order of magnitude higher. This suggests that dynein spends most of its time in the both bound state.

\section{State correlations}
If the both bound time is significantly higher than the one bound time, perhaps the two states are uncorrelated. In other words, does our model predict that the configuration of dynein before a step is unrelated to the final configuration when it rebinds again? To investigate this possible result, we plot the initial displacement versus the step length for the steps taken during all of the simulations. We consider these quantities as signed distances. For the step length, this indicates whether or not the model steps forwards or backwards. For the initial displacement, the sign indicates which binding domain initiates the step as indicated in the following figure. 
\begin{figure}[!hbt]
	\centering
	\includegraphics[angle=-90, width=1\columnwidth]{leading_vs_lagging}
	\caption[Leading vs Lagging Steps]{Diagram illustrating a leading foot step (a) and a lagging foot step (b).}
\end{figure}

\begin{figure}[!hbt]
	\centering
	\includegraphics[width=0.75\columnwidth]{paper_displacement_vs_step_length}
	\caption[Initial displacement vs Step Length]{\textbf{Two dimensional histogram} comparing the total signed step length to
			the signed initial displacement between the binding domains. A positive value
			indicates that the front foot made the step (leading). A negative value
			indicates the rear foot (lagging) made the step. Simulations show a strong,
			negative correlation. A similar fit from an experimental paper is plotted for comparison.}
	\label{fig:displacement_vs_step_length}
\end{figure}

Figure \ref{fig:displacement_vs_step_length} is a two-dimensional histogram for which step lengths are binned against initial displacements. A linear fit through the data (orange dashed line) indicates that the distribution is shifted vertically. This reflects the model's preference for stepping in the positive direction. Interestingly though, the slope of this fit appears much stepper than the same fit for experimental data from \cite{dewitt2012cytoplasmic}.

If we define the final step length to be 
\begin{equation}
\text{final displacement} = \text{initial displacement} + \text{step length },
\end{equation}	
then clearly our model indicates that the final displacement is roughly the same regardless of the initial displacement-- a conclusion at odds with the experimental data. This discrepancy may be due to the low number of data points in \cite{dewitt2012cytoplasmic}, but it can not be explained away by the expected lack of zero nm step observations.

\section{Simulating states separately}
Our simulations appear to predict that the one bound and both bound states are uncorrelated with each other. Therefore, one reasonable conclusion is: Why simulate them together in the first place? \\

We are now developing a simplified simulation code for our model which still performs the full Brownian dynamics for the short lived one bound state but takes advantage of this result by removing the both bound portion of the simulation all together. Instead, we seek to use Monte Carlo methods to generate both bound stepping statistics by determining unbinding based on the distribution of possible both bound configurations for a given initial displacement. \\

\begin{figure}[!hbt]
	\centering
	\includegraphics[width=1\columnwidth]{energy_landscape}
	\caption[Energy Landscape]{\textbf{Energy Landscape} graph showing the total energy of the dynein for a given binding domain displacement $L$. The yellow and red dots correspond to the minimum and maximum energy configurations for which a cartoon is shown on the right.}
	\label{fig:energy_landscape}
\end{figure}
In figure \ref{fig:energy_landscape} we can see a contour plot of the energy associated with dynein in the both bound state for all possible allowed combinations of near motor and far motor angles given an interhead separation of $16$ nm. Indicated on the graph are the minimum and maximum energy configurations (orange and red respectfully). The total energy appears to increase monotonically outwards from the minimum suggesting that the orange configuration shown to the right of the figure is a reasonable representative for the typical both bound state with this interhead separation.

We anticipate that a simulation distinguishing between the two states will significantly improve the computation time and enable us to fit simulation parameters more tactfully.

### 2019-05-07 papers/waczak-thesis/thesis/thesis.tex
thesis update 2
make thesis compile

\documentclass[12pt]{report}
\usepackage{graphicx} % make it easy to load in images
\graphicspath{{./images/},{../../../plots/}} % where to find images
\usepackage[a4paper, margin=1in]{geometry} % change margins 
\usepackage{fancyhdr} % add a header / footer 
\pagestyle{fancy}

@@

\title{
	{Simulating the Brownian Dynamics of a Two-dimensional Model for the Dynein Motor Protein}\vspace{1.5em}\\
	\includegraphics[width=0.5\columnwidth]{OSU_logo.png}\vspace{0.5em}\\
	\large{Oregon State University}
}
\author{John Waczak}
\date{\today}

@@

	\maketitle
	\todo[inline]{TODO: Fix formatting of title page}
	\chapter*{Abstract}
	\vspace{-3 em}word count: \textbf{345}\\

	The dynein molecule is a peculiar motor protein recognized for its odd stepping behavior. Sometimes it steps forwards. Other times it steps backwards. It has even been observed to occasionally shuffle by beginning a step with the same domain many times in row. These unique motions make dynein an interesting object of study, but its status as one of a select few minus-end directed motors makes understanding the mechanics of dynein’s motion vital to our understanding of the cell. When these motors misbehave, fundamental processes like axonal transport and muscle contraction fail. Observational techniques for studying biomolecules have revealed much about the different conformational states achieved by dynein during its stepping cycle and have established a base of relevant statistics such step lengths, velocities, and generated forces among others. Despite these advancements, the ability to image the full molecule for an entire stepping cycle remains beyond the capabilities of the current experimental apparatus.

	To address this gap, we have developed a simplified, two-dimensional model for the dynein molecule. We treat the protein as a system of massive domains held together by rigid rods steered towards equilibrium configurations by spring like torques and time evolved using Brownian dynamics. With our current set of parameters, the model displays directed motion with step length and velocity distributions ( give statistics here ) in agreement with experiment. Our simulations also suggest that there is no correlation between the conformation of dynein before and after a step. Our results indicate that a more computationally efficient simulation can be achieved with a two state system where a post power stroke state is selected from a distribution via a Monte Carlo simulation and binding is then simulated with the full Brownian dynamics. We are optimistic that our simulations will be able to address the power stroke versus winch model debate.

	\todo[inline]{TODO: Add in specific data to results}
	\todo[inline]{TODO: Fix formatting to include "a thesis submitted in completion of..." etc.}

	\chapter*{Acknowledgments}
	I want to thank ...

	\tableofcontents

	\chapter{Introduction}
	\input{./chapters/introduction}
	\chapter{Dynein}
	\input{./chapters/dynein}
	\chapter{Theory}
	\input{./chapters/theory}
	\chapter{Methods}
	\input{./chapters/methods}
	\chapter{Results}
	\input{./chapters/results}
\bibliography{references}
\bibliographystyle{prl_like.bst}
\end{document}

>>>>

\documentclass[12pt]{report}
\usepackage{graphicx} % make it easy to load in images
\graphicspath{{./images/},{../../../plots/}} % where to find images
\usepackage[letterpaper, margin=1in]{geometry} % change margins 
\usepackage{fancyhdr} % add a header / footer 
\pagestyle{fancy}

@@

\title{
	\large{Drunken Walking: A Novel Brownian Dynamics Simulation for the Mechanochemical Cycle of the Dynein Motor Protein}\vspace{1.5em}\\
	\normalsize{by}\\
	\large{John Waczak} \vspace{2em}\\ 
	\normalsize{an undergraduate thesis advised by David Roundy submitted to \\ 
		Department of Physics, Oregon State University}\vspace{2em}\\
	\normalsize{in partial fulfillment of the requirements for the degree of}\\
	\normalsize{Baccalaureate of Science in Physics}\vspace{1.5 em}\\
	\normalsize{Submitted \today}\vspace{3em}\\
	\includegraphics[width=0.5\columnwidth]{OSU_logo.png}\vspace{0.5em}\\
}
\date{}

@@

	\maketitle
	\chapter*{Abstract}
	\vspace{-3 em}word count: \textbf{282}\\

	The dynein molecule is a peculiar motor protein recognized for its odd stepping behavior. Sometimes it steps forwards. Other times it steps backwards. It has even been observed to occasionally shuffle by beginning a step with the same domain many times in row. These unique motions make dynein an interesting object of study, but its status as one of a select few minus-end directed motors makes understanding the mechanics of dynein’s motion vital to our understanding of the cell. When these motors misbehave, fundamental processes like axonal transport and muscle contraction fail. Experimental techniques for studying biomolecules have revealed much about the different conformational states achieved by dynein during its stepping cycle and have established a base of relevant statistics such step lengths, velocities, and generated forces among others. Despite these advancements, the ability to image the full molecule for an entire stepping cycle remains beyond current experimental capabilities.

	To address this gap, we have developed a simplified, two-dimensional model for the dynein molecule. We treat the protein as a system of massive domains held together by rigid rods steered towards equilibrium configurations by spring like torques and time evolved using Brownian dynamics. With our current set of parameters, the model displays directed motion with step length and velocity distributions in agreement with experiment. Our simulations also suggest that there is no correlation between the conformation of dynein before and after it takes a step. Our results indicate that a more computationally efficient simulation can be achieved with a two state system where a post power stroke state is selected from a distribution via a Monte Carlo simulation and binding is then simulated with the full Brownian dynamics.

	\tableofcontents
	\listoffigures
	%\listoftables

	\chapter{Introduction}
	\input{./chapters/introduction}
	\chapter{Dynein}
	\input{./chapters/dynein}
	\chapter{Theory}
	\input{./chapters/theory}
	\chapter{Methods}
	\input{./chapters/methods}
	\chapter{Results \& Discussion}
	\input{./chapters/results}
	\chapter{Conclusion}
	\input{./chapters/conclusion}
\bibliography{references}
\bibliographystyle{plain}
\end{document}
